Machine learning is a branch of computer science that seeks to use example data and past experience to improve algorithm performance~\cite{alpaydin:2004:introduction}. In this paper machine learning methods are evaluated using the Vinho Verde dataset~\cite{cortez:2009:modeling}. More specifically various machine learning methods are used to perform classification and regression tasks. 

Machine learning is a broad subject of research and its applications have garnered a lot of interest from the industry in recent years. This paper does not seek to provide in-depth analysis of the best possible methods on the given tasks but rather provide an experiment backed overview to couple of popular methods used in modern machine learning. A good overview to the basic machine learning methods and their mathematical grounding is given by Alpaydin in~\cite{alpaydin:2004:introduction}. The Alpaydin book covers most of the methods used in this paper except for neural networks based extreme learning machines. Neural networks are discussed in~\cite{haykin:2009:neural-networks}.

Experiments are conducted in this paper to demonstrate the performance and ease of use of different machine learning algorithms. The algorithms for this paper were selected with two goals in mind. The first goal was to find well performing algorithms for the course competition. The support vector machines and tree bagging were selected purely for their good performance in the dataset. The second goal was to select couple of machine learning algorithms that are both used in real world applications and easy to implement. K-nearest neighbor classifier, linear regression and extreme learning machines were selected for this purpose.

Neural networks have gained a lot of popularity lately as they have performed well in many real world machine learning problems as well as in widely used benchmarks such as the MNIST dataset~\cite{wan:2013:regularization, ciresan:2012:multi, haykin:2009:neural-networks}. Extreme learning machine was chosen to represent neural networks in this task as it is easy to implement and is based on similar mathematical model as the best performing neural networks.

The paper is laid out in five sections plus appendices. This introduction is the first section after which the methods used in this paper are explained in section~\ref{sec:methods}. Description of the conducted experiments follows in section~\ref{sec:experiments}. After the experiment decription the results are given in section~\ref{sec:results} and discussed in section~\ref{sec:discussion}. The algorithm implementations are given in the appendices.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "report"
%%% End:
